\documentclass[12pt]{article}

\usepackage{hyperref, tabularx, booktabs, amssymb,amsmath,amsfonts,eurosym,geometry,ulem,graphicx,caption,color,setspace,sectsty,comment,footmisc,caption,natbib,pdflscape,subfigure,array,hyperref, float}


\hypersetup{
    colorlinks=true,
    linkcolor=blue,
    filecolor=magenta,      
    urlcolor=cyan,
    pdftitle={Overleaf Example},
    pdfpagemode=FullScreen,
    }

\normalem

\onehalfspacing



\newcolumntype{L}[1]{>{\raggedright\let\newline\\arraybackslash\hspace{0pt}}m{#1}}
\newcolumntype{C}[1]{>{\centering\let\newline\\arraybackslash\hspace{0pt}}m{#1}}
\newcolumntype{R}[1]{>{\raggedleft\let\newline\\arraybackslash\hspace{0pt}}m{#1}}

\geometry{left=1.0in,right=1.0in,top=1.0in,bottom=1.0in}

\begin{document}

\begin{titlepage}
\title{Evictions, Education, and Identification Under Selection}
\author{Arjun Shanmugam}
\date{\today}
\maketitle
\begin{abstract}
\noindent I attempt to identify the causal effect of changes in the eviction rate on students' math and reading proficiency rates at the city level. Using data from Ohio and a fixed effects model, I estimate that a 1 percent increase in a city's eviction rate leads to a 0.064 percent decrease in its math proficiency rate. I find negative but insignificant effects on reading proficiency. In racially diverse cities, estimates are larger in magnitude and more significant. My fixed effects model controls for unobserved confounders that vary solely across cities or solely across time. But if unobserved confounders vary across both of these dimensions, then my estimates may be biased. I evaluate the plausibility of this concern by discussing eviction and its correlates and with an analysis of coefficient and $R^2$ stability when controls are introduced.\\

\bigskip
\end{abstract}
\setcounter{page}{0}
\thispagestyle{empty}
\end{titlepage}
\pagebreak \newpage




\doublespacing


\section{Introduction} \label{sec:introduction}
Fueled by stagnating incomes and rising housing costs, an eviction crisis has burdened low-income families in the United States for decades. It is estimated that over half of poor renting families spends over half its income on rent, meaning that unanticipated shocks can force families into homelessness \citep{desmond_evicted:_2017}. At various points beginning in March 2020, the federal government, the Centers for Disease Control, and some state and local governments implemented eviction moratoria, preventing a wave of evictions as pandemic-related lockdowns spread\citep{thrush_federal_2021, goldstein_landlords_2020}. But today, as moratoria expire, eviction rates are creeping up in cities across the country\citep{zaveri_after_2022}. A large body of research has investigated the effect of eviction on adults, but comparatively little evidence exists on how it affects children. If eviction impacts academic performance, then lifetime outcomes could be at stake.

This paper attempts to identify the causal effect of eviction on math and reading proficiency among third through eighth grade students. Ideally, longitudinal microdata would allow me to compare the test scores of children who have experienced eviction with the test scores of those who have not. Lacking this data, I am faced with the choice of conducting my analysis at the school district level or the city level. If eviction reduces test scores, and evicted students tend to move to different districts upon eviction, then an analysis at the school district level might underestimate the true effect of eviction. If students tend to attend school within the same \textit{city} upon being evicted, any change in their academic performance should be visible in city level data. Thus, I aggregate education and eviction data to the city level.

I draw the majority of my data from two sources. The first is The Eviction Lab at Princeton University, which has compiled a database of evicton rates over the 2000s and 2010s in nearly every city in the United States\citep{desmond_eviction_2018}. These data come from formal eviction records held by a number of public and private sector sources. The Eviction Lab also provides ACS 5-year estimates of time-varying socioeconomic characteristics of many of the cities in which it tracks eviction rates, such as the poverty rate and the percent homes which are renter occupied. My second source of data is the Ohio Department of Education, which provides district-level data on the portion of students judged proficient in math and reading from the 2005-06 school year to the 2014-2015 school year. In order to expand my set of controls, I also obtain ACS 5-year estimates of educational attainment at the city level. I build a novel panel dataset of eviction rates, math and reading proficiency scores, and a host of time-varying socioeconomic characteristics at the city-year level.

My biggest econometric challenge is that the treatment I seek to evaluate—eviction—is nonrandom. Partly due to generations of housing policies which sought to systematically exclude African-Americans from home ownership, eviction is strongly correlated with race \citep{rothstein_color_2017}. Eviction is also correlated with income, renter-occupied rates, and a host of other socioeconomic variables, many of which are also correlated with test scores. I address the threat to identification posed by observable confounders with a comprehensive set of controls.

Still, without a random source of variation in eviction rates, it is likely that eviction rates are correlated with unobserved characteristics of cities \footnote{Using \cite{kroeger_nuisance_2020}'s result that nuisance ordinances increase eviction rates, I attempted to instrument eviction rates using the presence of a nuisance ordinance in a city-year. Ultimately, the first-stage was too weak to produce precise instrumental variable estimates; I continue in search of an exogenous source of variation. I present the results of the first-stage regressions in the appendix.}. Using year and city fixed effects, I am able to control for unobserved, city-specific characteristics which are constant over time and unobserved, year-specific characteristics which are constant across cities. But time-variant unobservables—say, changes in landlords' attitudes towards tenants within a single city over time—may still bias my estimates. I do not claim to overcome this threat to identification. However, in Section 5, I argue that observed characteristics explain most of the variation in eviction rates and bound the bias in my estimated coefficients following \cite{oster_unobservable_2019}.

I contribute to a wide literature which studies the effects of eviction on important determinants of well being. Evicted mothers are more likely to be depressed; low-income workers are more likely to lose their jobs after being evicted; and at the height of the pandemic, eviction moratoria limited households’ food insecurity and mental stress\citep{desmond_housing_2016, desmond_evictions_2015, an_covid-19_2021}. Less research has focused on the effects of evictions on children. \cite{grigg_school_2012} finds that residence changes negatively affect outcomes such as academic achievement and social development, but evictions are traumatic for children in ways that voluntary moves are not. For this reason, and because education is central to the development of children, I hope to estimate the causal effect of eviction on test scores.

Section 2 discusses in greater depth the data I obtain and the dataset I assemble for my analysis. Section 3 outlines my empirical strategy. Section 4 provides and discusses results, and Section 6 concludes.
















\section{Background and Data} \label{sec:data}
The data used in this analysis come from three sources. I limit my analysis to the state of Ohio for two reasons. First, variation in eviction policy is relatively small across cities in Ohio (\textbf{CITE}). Second, by omitting other states from the analysis, I avoid the possibility that time and state-varying confounders—which state fixed effects cannot control for—bias my estimates.
\subsection{Eviction Data}
Eviction cases in America are usually tried in civil courts at the county-level. Most cases are resolved in one of three ways. An \textit{eviction judgement} can be rendered, "ordering the defendants to vacate a premise by a specific date"; the case can be \textit{dismissed}, "or ruled in favor of the defendant"; or the judge can permit a \textit{mediated agreement} between the landlord and the tenant, establishing certain terms of payment and resulting in an eviction judgement if these terms are unmet. In all three cases, courts generally keep records of evictions that include information such as landlord and tenant names and addresses, case rulings, and the date the case was heard.

I obtain eviction data from The Eviction Lab, which has collected millions of the above described records in two ways: first, by sending requests to state and county governments, and second, by purchasing records from two private companies. State and county records are publicly accessible under the Freedom of Information Act; about 12 million records were obtained in this manner. LexisNexis Risk Solutions and American Information Research Services, Inc. provided approximately 67 million and 11 million eviction records respectively. The Eviction Lab geocoded tenant addresses so that they could be linked to U.S. Census-designated cities, then aggregated records to the city-level to protect privacy and calculate eviction rates for each city during almost every year in the 2000s and 2010s. The Eviction Lab also merged socioeconomic variables taken from the American Community Survey's 2010 and 2015 five-year estimates: racial composition variables, the percent of households which are renter occupied, poverty rate, median household income, and median property value.

\subsection{Testing Data}

I obtain educational testing data from the Ohio Department of Education. The Ohio Department of Education implemented its current testing regime, the Ohio Achievement Assessment (OAA), soon after the passage of the No Child Left Behind Act, which required states to assess 3rd through 8th grade students in reading and math \citep{fox_examination_2014}. In 2010, the educational benchmarks used by the OAA were updated to incorporate the Common Core State Standards—a set of academic standards meant to ensure that all state curricula reached the same level of rigor \citep{ohio_school_boards_association_ohios_2010}.

The Ohio Department of Education has made OAA proficiency rates in math and reading at the school district-grade level available since the 2005-06 school year. I assemble this data into a single, school district-school year level dataset. Assessments are given in the spring, so I denote each school year by the year of its spring term (i.e., the 2005-06 school year is listed as year 2006). Using the National Center for Education Statistics (NCES) geographic relationship files, I match each school district to the city or cities that it intersects with. When a school district spans two cities, I calculate its \textit{adjusted enrollment} by dividing its total enrollment in each year equally across the cities it spans. Then, I aggregate the data to the city-year level. I calculate each city-year's proficiency rate in each subject as the mean of proficiency rates in the school districts spanned by the city, weighted by adjusted enrollment.

I then merge each city-year from the eviction data with the following city-year from the education data. The OAA is generally administered in May of each year. If I were to match on identical city-years, I would risk attempting to draw a causal link between evictions and \textit{past} standardized test performance.



\section{Empirical Strategy} \label{sec:empirical_strategy}

\section{Results and Discussion} \label{sec:result}

\section{Conclusion} \label{sec:conclusion}



\singlespacing
\setlength\bibsep{0pt}
\bibliographystyle{rusnat}
\bibliography{citations}



\clearpage

\onehalfspacing

\section*{Tables} \label{sec:tab}
\addcontentsline{toc}{section}{Tables}
\input{output/tables/summary_stats}
\input{output/tables/main_regressions}
\input{output/tables/diverse_regressions}
\input{output/tables/non_diverse_regressions}


\clearpage

\section*{Figures} \label{sec:fig}
\addcontentsline{toc}{section}{Figures}



\clearpage

\section*{Appendix A. Placeholder} \label{sec:appendixa}
first stage results from attempted instrumental variables approach
\addcontentsline{toc}{section}{Appendix A}



\end{document}